---
title: "Homework 6"
author: "Ellen Chancey"
date: "November 18, 2017"
output: html_document
---
## Chapter 7 Project 7.38 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/EC052367/Documents/MS-ASA/STAT 840 Linear Regression/Homework/Lesson 6")

# clear the environment
rm(list=ls())

C1<- read.csv("C1 SENIC.csv")
# summary(C1)

# install packages
library("Hmisc")
library("rgl")
library("rsq")
library("dplyr")
```

Assigning variables
```{r}
# set variables
y <- C1$los
x1 <- C1$age
x2 <- C1$ir
x3 <- C1$rcr
x4 <- C1$adc
x5 <- C1$nn
x6 <- C1$afs
```

Initial Model Views
```{r}
plot3d(x1,y,x2,type="h") # initial 3d view
points3d(x1,y,x2,col="red") # makes dot red
```

```{r}
pairs(y~x1+x2)
```

Model Set Up
```{r}
# model set up
m1 <- lm(y~x1+x2)
summary(m1)
```

Setup Alternative Models
```{r}
# note that model is one number below the added variable
# m1 acts as the reduced model
m2 <- lm(y~x1+x2+x3)
m3 <- lm(y~x1+x2+x4)
m4 <- lm(y~x1+x2+x5)
m5 <- lm(y~x1+x2+x6)
```


#### A) Calculating the Coefficient of Partial Determination
```{r}
rsq1 <- rsq.partial(m1,m2)
rsq2 <- rsq.partial(m1,m3)
rsq3 <- rsq.partial(m1,m4)
rsq4 <- rsq.partial(m1,m5)
cpd <- c(rsq1$partial.rsq, rsq2$partial.rsq, rsq3$partial.rsq, rsq4$partial.rsq)
models <- c(2:5)
cpd_all <- cbind(models,cpd)
cpd_all

```

#### B) Routine Culture Ratio  
Model three reduces the error by the largest amount. Model two introduces $x_4$, average daily census.  


#### C) Testing Routine Culturing Ratio
$H_0: \beta_4 = 0$  
$H_1: \beta_4 <> 0$
```{r}
# compare model one and three (with and without x_4)
anova2 <- anova(m1,m3)
p<- anova2$`Pr(>F)`[2]
anova2
```

#### F Test
The p value `r p` is sufficient to reject the null Hypothesis, meaning $\beta_4$ is not equal to zero and can improve the model.

#### F Test of Other Xs
```{r}
anova1 <- anova(m1,m2)
anova2 <- anova(m1,m3)
anova3 <- anova(m1,m4)
anova4 <- anova(m1,m5)

f_all <- c(anova1$F[2],anova2$F[2],anova3$F[2],anova4$F[2])
p_all <- c(anova1$`Pr(>F)`[2],anova2$`Pr(>F)`[2],anova3$`Pr(>F)`[2],anova4$`Pr(>F)`[2])
f_p_all<- cbind(models,f_all,p_all)
f_p_all
```

#### Comparing F tests of other variables  
Models four and five, which add number of nurses and available facilities and services, respectively, would also improve the model. These would not improve as much as model three. Model two does not reject the null hypothesis, and therefore would not add anything to the model.

```{r session-info}
sessionInfo()
```